Informing IPCC accounting of forest carbon using the global forest carbon database (ForC v4.0)

### 2022-04-27 doc/manuscript/ESSDmanuscript.Rmd
reposition table
add in-prep note
improve table formatting

    header-includes:
    - \usepackage{caption}
---

@@

@@

library(knitr)
library(kableExtra)
knitr::kable(C_pools_table, format = "latex",
             longtable = FALSE,
             caption = "\\textbf{IPCC-defined forest carbon pools with definitions and measurement methods.} Definitions from IPCC Table 1.1. (See Table 1.1 in IPCC guidance).") %>%
  kableExtra::kable_styling(full_width = TRUE) %>%
  #column_spec(1, width = ".7cm") %>% #This doesn't seem to work
  row_spec(0, bold = TRUE)
```

@@

```{r table_variables, eval = TRUE, echo=FALSE, warning=FALSE}
C_variables_table <- read.csv("figures_tables/C_variables.csv")
names(C_variables_table) <- stringr::str_replace_all(names(C_variables_table), "\\.", " ")
C_variables_table[is.na(C_variables_table)]<- "" #replace NAs with blanks

library(knitr)
library(kableExtra)
knitr::kable(C_variables_table[,-1], format = "latex",
             longtable = TRUE,
             caption = "\\textbf{Numbers of records of ForC variables relevant to, and sent to, EFDB.} Relationships of variables to IPCC-defined forest C pools (Table 1) and to each other are illustrated in Figure 1. ") %>%
  kableExtra::kable_styling(full_width = TRUE) %>%
  row_spec(0, bold = TRUE)  %>%
  row_spec(which(C_variables_table$bold), bold = TRUE) #total
```

@@

@@

library(knitr)
library(kableExtra)
knitr::kable(recommendations_table, format = "latex",
             longtable = FALSE,
             caption = "\\textbf{Recommended best practices for reporting forest C estimates of value to national greenhouse gas inventories under IPCC guidance. } ....") %>%
  kableExtra::kable_styling(full_width = TRUE) %>%
  row_spec(0, bold = TRUE)
```

@@

library(knitr)
library(kableExtra)
knitr::kable(record_of_changes, format = "latex",
             longtable = TRUE) %>%
  kableExtra::kable_styling(full_width = TRUE, latex_options = c("repeat_header")) %>%
  row_spec(0, bold = TRUE)
```

@@

knitr::kable(ForC_variables_included, format = "latex",
             longtable = TRUE) %>%
  row_spec(0, bold = TRUE) %>%
kableExtra::kable_styling(full_width = TRUE, latex_options = c("repeat_header", "hold position"))

>>>>

    header-includes:
    - \usepackage{caption}
    #- \usepackage{booktabs}
    #- \usepackage{makecell}
---

@@

****
*THIS IS AN IN-PREP MANUSCRIPT.*

****

@@

library(knitr)
library(kableExtra)
knitr::kable(C_pools_table, format = "latex",
             longtable = FALSE,
             caption = "\\textbf{IPCC-defined forest carbon pools with definitions and measurement methods.} Definitions from IPCC Table 1.1. (See Table 1.1 in IPCC guidance).") %>%
  row_spec(0, bold = TRUE) %>%
  kableExtra::kable_styling(full_width = TRUE) 
  #kable_styling(latex_options = c("scale_down", "hold_position"), protect_latex = T) 
```

@@

\newpage
```{r table_variables, eval = TRUE, echo=FALSE, warning=FALSE}
C_variables_table <- read.csv("figures_tables/C_variables.csv")
names(C_variables_table) <- stringr::str_replace_all(names(C_variables_table), "\\.", " ")
C_variables_table[is.na(C_variables_table)]<- "" #replace NAs with blanks

library(knitr)
library(kableExtra)
knitr::kable(C_variables_table[,-1], format = "latex",  
             longtable = T,
             caption = "\\textbf{Numbers of records of ForC variables relevant to, and sent to, EFDB.} Relationships of variables to IPCC-defined forest C pools (Table 1) and to each other are illustrated in Figure 1. ") %>%
  row_spec(0, bold = TRUE)  %>%
  row_spec(which(C_variables_table$bold), bold = TRUE) %>% #total
  kable_styling(font_size = 8) %>%
  #kableExtra::kable_styling(full_width = TRUE) 
  kable_styling(latex_options = c("hold_position", "repeat_header"), protect_latex = T) 
```

@@

library(knitr)
library(kableExtra)
knitr::kable(recommendations_table, format = "latex",
             longtable = FALSE,
             caption = "\\textbf{Recommended best practices for reporting forest C estimates of value to national greenhouse gas inventories under IPCC guidance. } ....") %>%
  row_spec(0, bold = TRUE)%>%
  kable_styling(font_size = 10) %>%
  kableExtra::kable_styling(full_width = TRUE) 
  #kable_styling(latex_options = c("scale_down", "hold_position"), protect_latex = T) 
```

@@

@@

library(knitr)
library(kableExtra)
knitr::kable(record_of_changes, format = "latex",
             longtable = TRUE) %>%
  kableExtra::kable_styling(full_width = TRUE, latex_options = c("repeat_header")) %>%
  kable_styling(font_size = 8) %>%
  row_spec(0, bold = TRUE)
```

@@

knitr::kable(ForC_variables_included, format = "latex",
             longtable = TRUE) %>%
  row_spec(0, bold = TRUE) %>%
  kable_styling(font_size = 8) %>%
kableExtra::kable_styling(full_width = TRUE, latex_options = c("repeat_header", "hold position"))

### 2022-04-27 doc/manuscript/ESSDmanuscript.tex
improve table formatting

To transfer complete, reviewed ForC records into EFDB, we created R
scripts to restructure ForC records and populate EFDB's bulk import form
(``EFDB bulk import.xlsx''). Criteria for data transfer were that (1)
records had been checked against the original study and determined to be
complete and correct (\texttt{EFDB.ready} field in
\texttt{ForC\_citations} table), (2) the original study presented values
in tables or text, as opposed to the values having been digitized from
graphs or calculated based on related variables
(\texttt{data.location.within.source} field in
\texttt{ForC\_measurements} table), and (3) the records had not
previously been sent to EFDB. Once converted into EFDB format, the
records were reviewed and then sent to the IPCC's Technical Support Unit
for inclusion in EFDB. Complete records needed to be reviewed by the
EFDB editorial board prior to posting in the database -- a process that
lags behind records transfer and had not yet been completed for all
records sent as of April 26, 2022.

@@

As of April 26, 2022, ForC (v4.0) contained 32550 independent records
(39856 total), 17190 of which were for variables relevant to EFDB (Fig.
1). These records were distributed across all forested continents and
ecozones, with particularly high concentrations in
\emph{{[}ecozone(s){]}} and low concentrations in
\emph{{[}ecozone(s){]}} (Fig. 2). The most widely represented forest
type was \emph{{[}type{]}}, followed by \emph{{[}type{]}} and
\emph{{[}type{]}} (Fig. 3).

@@

\begin{longtabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\caption{\label{tab:table_variables}\textbf{Numbers of records of ForC variables relevant to, and sent to, EFDB.} Relationships of variables to IPCC-defined forest C pools (Table 1) and to each other are illustrated in Figure 1. }\\
\hline
\textbf{variable} & \textbf{n in ForC} & \textbf{n independent records in ForC} & \textbf{n reviewed} & \textbf{n sent to EFDB}\\
\hline
\textbf{Biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass & 1095 & 847 & 93 & 48\\
\hline
delta.biomass & 0 & 0 & 0 & 0\\
\hline
NPP\_woody & 136 & 93 & 0 & 0\\
\hline
woody.mortality & 0 & 0 & 0 & 0\\
\hline
\textbf{Aboveground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_ag & 9038 & 7724 & 1232 & 693\\
\hline
biomass\_ag\_woody & 460 & 366 & 10 & 10\\
\hline
biomass\_ag\_foliage & 601 & 502 & 49 & 27\\
\hline
delta.agb & 166 & 128 & 123 & 123\\
\hline
ANPP\_woody & 299 & 242 & 0 & 0\\
\hline
ANPP\_foliage & 0 & 0 & 0 & 0\\
\hline
woody.mortality\_ag & 113 & 61 & 17 & 17\\
\hline
\textbf{Belowground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_root & 4629 & 4180 & 123 & 55\\
\hline
biomass\_root\_fine & 931 & 594 & 18 & 18\\
\hline
biomass\_root\_coarse & 599 & 410 & 12 & 7\\
\hline
delta.biomass\_root & 0 & 0 & 0 & 0\\
\hline
delta.biomass\_root\_coarse & 0 & 0 & 0 & 0\\
\hline
delta.biomass\_root\_fine & 0 & 0 & 0 & 0\\
\hline
woody.mortality\_root & 0 & 0 & 0 & 0\\
\hline
BNPP\_root\_fine & 489 & 333 & 0 & 0\\
\hline
BNPP\_root.turnover\_fine & 91 & 56 & 0 & 0\\
\hline
BNPP\_root\_coarse & 329 & 250 & 0 & 0\\
\hline
\textbf{Dead wood} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
deadwood & 438 & 303 & 94 & 61\\
\hline
deadwood\_standing & 152 & 120 & 17 & 17\\
\hline
deadwood\_down & 424 & 368 & 51 & 27\\
\hline
delta.deadwood & 0 & 0 & 0 & 0\\
\hline
delta.deadwood\_standing & 0 & 0 & 0 & 0\\
\hline
delta.deadwood\_down & 0 & 0 & 0 & 0\\
\hline
R\_het\_deadwood & 0 & 0 & 0 & 0\\
\hline
\textbf{Litter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
O.horizon & 38 & 38 & 38 & 38\\
\hline
delta.O.horizon & 4 & 4 & 4 & 4\\
\hline
litter & 30 & 30 & 23 & 23\\
\hline
delta.litter & 0 & 0 & 0 & 0\\
\hline
NPP\_litter & 0 & 0 & 0 & 0\\
\hline
R\_het\_litter & 167 & 143 & 0 & 0\\
\hline
\textbf{Soil organic matter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
SOM / SOC & 693 & 398 & 81 & 46\\
\hline
delta.SOM / delta.SOC & 0 & 0 & 0 & 0\\
\hline
R\_het\_soil & 0 & 0 & 0 & 0\\
\hline
\textbf{TOTAL} & \textbf{20922} & \textbf{17190} & \textbf{1985} & \textbf{1214}\\
\hline
\end{longtabu}

@@

As of April 26, 2022, we had reviewed or added 1985 EFDB-relevant
records, 1214 records of which were sent to EFDB, and 73 of which have
been reviewed, accepted, and posted (Figs. 2-3, Table 2).
\emph{{[}DETAILS{]}}

@@

\caption{\label{tab:table_recommendations}\textbf{Recommended best practices for reporting forest C estimates of value to national greenhouse gas inventories under IPCC guidance. } ....}
\centering
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\hline
\textbf{criteria} & \textbf{recommendation} & \textbf{rationale}\\
\hline
variables to include & When possible, calculate and present all relevant variables that can be readily caluclated based on available data. & Estimates of relevant variables are not always calcualted.\\
\hline
forest census methods & Adopt IPCC guidelines (country-specific) for minimum stem size in censues in census and reporting. Ideally, census stem down to the smallest diameter feasible. & IPCC biomass pool definition includes all living vegetation, but understory may be excluded when contribution is minor.\\
\hline
 & *taxa to include* & \\
\hline
dead organic matter sampling & *include damage estimates on standing dead trees* & \\
\hline
 & Adopt IPCC recommendations for minimum diameter of deadwood (country-specific, default 10 cm). & \\
\hline
belowground sampling & Select and report soil sampling increments to include a cutoff at 30 cm depth (or country-specific depth). & IPCC biomass pool definition includes all living vegetation\\
\hline
reporting variiables & Present each variable individually, as opposed to requiring that variables of interest be calculated from related variables. & EFDB requires that values in the database be presented in the original article, and cannot accept subsequent calculations.\\
\hline
reporting estimates & Report all relevant values in tables, text, or supplementary tables/ data files, as opposed to in figures only. & EFDB does not acceptvalues digitized from figures.\\
\hline
reporting confidence intervals & Report 95\% confidence intervals, standard error, or standard deviation and sample size. & EFDB requires confidence invervals, when possible.\\
\hline
\end{tabu}
\end{table}

@@

Table A1: \textbf{Table of changes to ForC fields.}

\begin{longtabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\hline
\textbf{Table} & \textbf{Column} & \textbf{Description} & \textbf{Changes} & \textbf{Motivation}\\
\hline
\endfirsthead
\multicolumn{5}{@{}l}{\textit{(continued)}}\\
\hline
\textbf{Table} & \textbf{Column} & \textbf{Description} & \textbf{Changes} & \textbf{Motivation}\\
\hline
\endhead
Sites & coordinates.precision & Precision of geographic coordinates, as reported by source or estimated from maps. & field added & allow identification of records with poor coordinate precision\\
\hline
Measurements & data.location.within.source & Location of data within the source listed in citation.ID. & field added & facilitate review, ensure traceability\\
\hline
 & sd, se, lower95\%CI, upper 95\%CI & Standard deviation, standard error, and lower and upper 95 percent confidence intvervals, respectively. & replaces `stat` and `stat.name` & cleaner format; ability to handle assymetrical 95 percent confidence intervals\\
\hline
 & mean.in.original.units, original.units & mean value and units presented in original publication & fields added & provide IPCC with original units, reduce errors/improve reproducibility\\
\hline
 & C.conversion.factor & Assumed/ measured C content of organic matter used to convert organic matter to C. & field added & track units conversion, allow back-calculation of OM if conversion factor deemed inappropriate\\
\hline
PFT & description & Definition of the pftcode at the community level. Differs from individual level in that properly describes mixed plant functional types. & field added & \\
\hline
 & description.individual & Definition of the pftcode at the individual plant level. & field name change (previously `description`) & \\
\hline
Citations & citation.citation & Full citation. Most of these records are automatically generated in R based upon DOI lookup. & field added & field required by IPCC\\
\hline
 & citation.language & Language of original publication, automatically generated based on the title and abstract, with some manual entries and corrections. & field added & field required by IPCC\\
\hline
 & citation.url & URL of original publication, generally retrieved automatically via URL lookup. & field added & field required by IPCC\\
\hline
 & citation.abstract & Abstract, generally retrieved automatically via DOI lookup. & field added & field required by IPCC\\
\hline
 & source.type & citation source type & field added & field required by IPCC\\
\hline
 & pdf.in.repository & Indicates whether pdf of original study has been retrieved and saved in ForC's reference repository & field added & housekeeping\\
\hline
 & EFDB.ready & Indicates whether data have been checked for export to EFDB. & field added & housekeeping\\
\hline
\end{longtabu}

@@

Table B1: \textbf{Mapping of ForC fields to EFDB.} See footnotes at end
of table (STILL NEED TO BE PROPERLY INSERTED).

@@

>>>>

To transfer complete, reviewed ForC records into EFDB, we created R
scripts to restructure ForC records and populate EFDB's bulk import form
(``EFDB bulk import.xlsx''). Criteria for data transfer were that (1)
records had been checked against the original study and determined to be
complete and correct (\texttt{EFDB.ready} field in
\texttt{ForC\_citations} table), (2) the original study presented values
in tables or text, as opposed to the values having been digitized from
graphs or calculated based on related variables
(\texttt{data.location.within.source} field in
\texttt{ForC\_measurements} table), and (3) the records had not
previously been sent to EFDB. Once converted into EFDB format, the
records were reviewed and then sent to the IPCC's Technical Support Unit
for inclusion in EFDB. Complete records needed to be reviewed by the
EFDB editorial board prior to posting in the database -- a process that
lags behind records transfer and had not yet been completed for all
records sent as of April 27, 2022.

@@

As of April 27, 2022, ForC (v4.0) contained 32550 independent records
(39856 total), 17190 of which were for variables relevant to EFDB (Fig.
1). These records were distributed across all forested continents and
ecozones, with particularly high concentrations in
\emph{{[}ecozone(s){]}} and low concentrations in
\emph{{[}ecozone(s){]}} (Fig. 2). The most widely represented forest
type was \emph{{[}type{]}}, followed by \emph{{[}type{]}} and
\emph{{[}type{]}} (Fig. 3).

@@

\newpage
\begingroup\fontsize{8}{10}\selectfont

\begin{longtable}[t]{l|l|l|l|l}
\caption{\label{tab:table_variables}\textbf{Numbers of records of ForC variables relevant to, and sent to, EFDB.} Relationships of variables to IPCC-defined forest C pools (Table 1) and to each other are illustrated in Figure 1. }\\
\hline
\textbf{variable} & \textbf{n in ForC} & \textbf{n independent records in ForC} & \textbf{n reviewed} & \textbf{n sent to EFDB}\\
\hline
\endfirsthead
\caption[]{\textbf{Numbers of records of ForC variables relevant to, and sent to, EFDB.} Relationships of variables to IPCC-defined forest C pools (Table 1) and to each other are illustrated in Figure 1.  \textit{(continued)}}\\
\hline
\textbf{variable} & \textbf{n in ForC} & \textbf{n independent records in ForC} & \textbf{n reviewed} & \textbf{n sent to EFDB}\\
\hline
\endhead
\textbf{Biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass & 1095 & 847 & 93 & 48\\
\hline
delta.biomass & 0 & 0 & 0 & 0\\
\hline
NPP\_woody & 136 & 93 & 0 & 0\\
\hline
woody.mortality & 0 & 0 & 0 & 0\\
\hline
\textbf{Aboveground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_ag & 9038 & 7724 & 1232 & 693\\
\hline
biomass\_ag\_woody & 460 & 366 & 10 & 10\\
\hline
biomass\_ag\_foliage & 601 & 502 & 49 & 27\\
\hline
delta.agb & 166 & 128 & 123 & 123\\
\hline
ANPP\_woody & 299 & 242 & 0 & 0\\
\hline
ANPP\_foliage & 0 & 0 & 0 & 0\\
\hline
woody.mortality\_ag & 113 & 61 & 17 & 17\\
\hline
\textbf{Belowground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_root & 4629 & 4180 & 123 & 55\\
\hline
biomass\_root\_fine & 931 & 594 & 18 & 18\\
\hline
biomass\_root\_coarse & 599 & 410 & 12 & 7\\
\hline
delta.biomass\_root & 0 & 0 & 0 & 0\\
\hline
delta.biomass\_root\_coarse & 0 & 0 & 0 & 0\\
\hline
delta.biomass\_root\_fine & 0 & 0 & 0 & 0\\
\hline
woody.mortality\_root & 0 & 0 & 0 & 0\\
\hline
BNPP\_root\_fine & 489 & 333 & 0 & 0\\
\hline
BNPP\_root.turnover\_fine & 91 & 56 & 0 & 0\\
\hline
BNPP\_root\_coarse & 329 & 250 & 0 & 0\\
\hline
\textbf{Dead wood} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
deadwood & 438 & 303 & 94 & 61\\
\hline
deadwood\_standing & 152 & 120 & 17 & 17\\
\hline
deadwood\_down & 424 & 368 & 51 & 27\\
\hline
delta.deadwood & 0 & 0 & 0 & 0\\
\hline
delta.deadwood\_standing & 0 & 0 & 0 & 0\\
\hline
delta.deadwood\_down & 0 & 0 & 0 & 0\\
\hline
R\_het\_deadwood & 0 & 0 & 0 & 0\\
\hline
\textbf{Litter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
O.horizon & 38 & 38 & 38 & 38\\
\hline
delta.O.horizon & 4 & 4 & 4 & 4\\
\hline
litter & 30 & 30 & 23 & 23\\
\hline
delta.litter & 0 & 0 & 0 & 0\\
\hline
NPP\_litter & 0 & 0 & 0 & 0\\
\hline
R\_het\_litter & 167 & 143 & 0 & 0\\
\hline
\textbf{Soil organic matter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
SOM / SOC & 693 & 398 & 81 & 46\\
\hline
delta.SOM / delta.SOC & 0 & 0 & 0 & 0\\
\hline
R\_het\_soil & 0 & 0 & 0 & 0\\
\hline
\textbf{TOTAL} & \textbf{20922} & \textbf{17190} & \textbf{1985} & \textbf{1214}\\
\hline
\end{longtable}
\endgroup{}

@@

As of April 27, 2022, we had reviewed or added 1985 EFDB-relevant
records, 1214 records of which were sent to EFDB, and 73 of which have
been reviewed, accepted, and posted (Figs. 2-3, Table 2).
\emph{{[}DETAILS{]}}

@@

\caption{\label{tab:table_recommendations}\textbf{Recommended best practices for reporting forest C estimates of value to national greenhouse gas inventories under IPCC guidance. } ....}
\centering
\fontsize{10}{12}\selectfont
\begin{tabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\hline
\textbf{criteria} & \textbf{recommendation} & \textbf{rationale}\\
\hline
variables to include & When possible, calculate and present all relevant variables that can be readily caluclated based on available data. & Estimates of relevant variables are not always calcualted.\\
\hline
forest census methods & Adopt IPCC guidelines (country-specific) for minimum stem size in censues in census and reporting. Ideally, census stem down to the smallest diameter feasible. & IPCC biomass pool definition includes all living vegetation, but understory may be excluded when contribution is minor.\\
\hline
 & *taxa to include* & \\
\hline
dead organic matter sampling & *include damage estimates on standing dead trees* & \\
\hline
 & Adopt IPCC recommendations for minimum diameter of deadwood (country-specific, default 10 cm). & \\
\hline
belowground sampling & Select and report soil sampling increments to include a cutoff at 30 cm depth (or country-specific depth). & IPCC biomass pool definition includes all living vegetation\\
\hline
reporting variiables & Present each variable individually, as opposed to requiring that variables of interest be calculated from related variables. & EFDB requires that values in the database be presented in the original article, and cannot accept subsequent calculations.\\
\hline
reporting estimates & Report all relevant values in tables, text, or supplementary tables/ data files, as opposed to in figures only. & EFDB does not acceptvalues digitized from figures.\\
\hline
reporting confidence intervals & Report 95\% confidence intervals, standard error, or standard deviation and sample size. & EFDB requires confidence invervals, when possible.\\
\hline
\end{tabu}
\end{table}

@@

Table A1: \textbf{Table of changes to ForC fields.}
\begingroup\fontsize{8}{10}\selectfont

\begin{longtabu} to \linewidth {>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X>{\raggedright}X}
\hline
\textbf{Table} & \textbf{Column} & \textbf{Description} & \textbf{Changes} & \textbf{Motivation}\\
\hline
\endfirsthead
\multicolumn{5}{@{}l}{\textit{(continued)}}\\
\hline
\textbf{Table} & \textbf{Column} & \textbf{Description} & \textbf{Changes} & \textbf{Motivation}\\
\hline
\endhead
Sites & coordinates.precision & Precision of geographic coordinates, as reported by source or estimated from maps. & field added & allow identification of records with poor coordinate precision\\
\hline
Measurements & data.location.within.source & Location of data within the source listed in citation.ID. & field added & facilitate review, ensure traceability\\
\hline
 & sd, se, lower95\%CI, upper 95\%CI & Standard deviation, standard error, and lower and upper 95 percent confidence intvervals, respectively. & replaces `stat` and `stat.name` & cleaner format; ability to handle assymetrical 95 percent confidence intervals\\
\hline
 & mean.in.original.units, original.units & mean value and units presented in original publication & fields added & provide IPCC with original units, reduce errors/improve reproducibility\\
\hline
 & C.conversion.factor & Assumed/ measured C content of organic matter used to convert organic matter to C. & field added & track units conversion, allow back-calculation of OM if conversion factor deemed inappropriate\\
\hline
PFT & description & Definition of the pftcode at the community level. Differs from individual level in that properly describes mixed plant functional types. & field added & \\
\hline
 & description.individual & Definition of the pftcode at the individual plant level. & field name change (previously `description`) & \\
\hline
Citations & citation.citation & Full citation. Most of these records are automatically generated in R based upon DOI lookup. & field added & field required by IPCC\\
\hline
 & citation.language & Language of original publication, automatically generated based on the title and abstract, with some manual entries and corrections. & field added & field required by IPCC\\
\hline
 & citation.url & URL of original publication, generally retrieved automatically via URL lookup. & field added & field required by IPCC\\
\hline
 & citation.abstract & Abstract, generally retrieved automatically via DOI lookup. & field added & field required by IPCC\\
\hline
 & source.type & citation source type & field added & field required by IPCC\\
\hline
 & pdf.in.repository & Indicates whether pdf of original study has been retrieved and saved in ForC's reference repository & field added & housekeeping\\
\hline
 & EFDB.ready & Indicates whether data have been checked for export to EFDB. & field added & housekeeping\\
\hline
\end{longtabu}
\endgroup{}

@@

Table B1: \textbf{Mapping of ForC fields to EFDB.} See footnotes at end
of table (STILL NEED TO BE PROPERLY INSERTED).
\begingroup\fontsize{8}{10}\selectfont

@@

\endgroup{}
